---
title: "Data Science I, Midterm Project"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r Packages, message=FALSE}
# Load Necessary Packages
  library(ggplot2)
  library(tidyverse)
  library(dplyr)
  library(tidyr)
  library(patchwork)

# Preliminary Loading of All Datasets
  NYC_Rental_Price = read.csv("./NYC Rental.csv")
  NYC_Zipcode      = read.csv("./NYC Zipcode.csv")
  US_Housing_Price = read.csv("./US Housing 2023.csv")

```

##### [Discussion] Description and Processing of Raw Datasets   
- `NYC_Rental_Price Dataset`    
  - Description: contain detailed NYC Rental Prices between Jan.2015 to Aug.2024 listed by zip codes   
  - Dimension: `r nrow(NYC_Rental_Price)` x `r ncol(NYC_Rental_Price)`    
  - Unique Zipcodes: `r n_distinct(NYC_Rental_Price$RegionName)`
  - Issues:   
    - ambiguous column names        
      - Column name `RegionName` actually consists entirely of zipcode information, which is very misleading.   
      - Column names of form `X2015.1.31` should be renamed for better readability.    
    - redundant columns   
      - Column `RegionID` is useless as we can treat the zipcode variable as the unique identifier.
      - The entire dataset is within NYC so columns `StateName`, `State`, `City`, `Metro` can be removed.
      - Column `SizeRank` is never used for future analysis.
      - Column `RegionType` has only `r n_distinct(NYC_Rental_Price$RegionType)` level of value and is useless.
    - too much missing data    
  - Data Cleaning Strategy:
    - Rename and Keep only the following columns: `county`, `zipcode`, and all date columns of form `%yyyy%mm%dd`.
    - Remove records containing empty values.
- `NYC_Zipcode Dataset`   
  - Description: contain detailed zipcode-location associations within NYC    
  - Dimension: `r nrow(NYC_Zipcode)` x `r ncol(NYC_Zipcode)`    
  - Unique Zipcodes: `r n_distinct(NYC_Zipcode$ZipCode)`
  - Issues:
    - redundant columns
      - Only column `County`, `ZipCode`, `Neighborhood` is useful.
  - Data Cleaning Strategy:
    - Rename and Keep only columns `county` and `zipcode`.
    - Also per instruction required, add `borough` variable.
- `US_Housing_Price Dataset`    
  - Description: contains detailed housing prices information within entire U.S. associated with specific area    
  - Dimension: `r nrow(US_Housing_Price)` x `r ncol(US_Housing_Price)`   
  - Issues:   
    - ambiguous columns   
      - Column name `RegionName` actually consists entirely of zipcode information, which is very misleading. 
      - Column names of form `X2015.1.31` should be renamed for better readability.  
    - redundant columns   
      - Only zipcodes, State, and date columns are used in future analysis.   
    - missing data
  - Data Cleaning Strategy:   
    - Rename and Keep only variables `zipcode`, `state`, and all date columns of form `%yyyy%mm%dd`.
    - Remove records containing empty values.

##### [Code Chunk] Raw Data Cleaning    
```{r Raw Data Cleaning}

  NYC_Rental_Price = NYC_Rental_Price %>% janitor::clean_names() %>% select(-region_id, -size_rank, -region_type, -state_name, -state, -city, -metro) %>% drop_na() %>%
                     rename("zipcode"="region_name", "county"="county_name") %>% mutate(county=gsub(" County","",county)) %>% mutate(across(where(is.numeric), \(x) round(x, 2))) %>% arrange(zipcode)
                     colnames(NYC_Rental_Price)[3:ncol(NYC_Rental_Price)] <- colnames(NYC_Rental_Price)[3:ncol(NYC_Rental_Price)] %>% sub("x", "", .) %>% gsub("_", "/", .) %>%
                                                                             sub("(\\d{4})/(\\d{1})/(\\d{2})", "\\1/0\\2/\\3", .)
                     
  NYC_Zipcode = NYC_Zipcode %>% janitor::clean_names() %>% select(county, zip_code, neighborhood) %>% rename("zipcode"="zip_code") %>% drop_na() %>%
                mutate(borough=recode(county, "Kings"="Brooklyn", "New York"="Manhattan", "Richmond"="Staten Island")) %>% relocate(zipcode, county, borough)
  
  US_Housing_Price = US_Housing_Price %>% janitor::clean_names() %>% select(region_name, state, starts_with("x")) %>% rename("zipcode"="region_name") %>% arrange(zipcode) %>% drop_na()
                     colnames(US_Housing_Price)[3:ncol(US_Housing_Price)] <- colnames(US_Housing_Price)[3:ncol(US_Housing_Price)] %>% sub("x", "", .) %>% gsub("_", "/", .) %>%
                                                                             sub("(\\d{4})/(\\d{1})/(\\d{2})", "\\1/0\\2/\\3", .)
                     
```

##### [Code Chunk] Merge NYC_Rental and NYC_Zipcode Datasets       
```{r Merge Datasets}

  Merged = left_join(x=NYC_Zipcode, y=NYC_Rental_Price, by="zipcode") %>% rename("county"="county.x") %>% select(-county.y) %>% relocate(zipcode, county, borough, neighborhood) %>% drop_na()

```
##### [Discussion] Merged Datasets, Issues, Fixing
- The merged dataset has several issues, firstly there are many missing values under the county column from the `NYC_Rental_Price` dataset, so we drop `county.y` as all information is
  available in `county.y`. The merged dataset has no missing values for the character variables `zipcode`, `borough`, `county` but has plenty of missing values for the numeric ones.
  Remove records containing empty values.

##### [Discussion] on TIDYed Dataset       

- `NYC_Rental_Price` Dataset
  - Dimension: `r nrow(NYC_Rental_Price)` x `r ncol(NYC_Rental_Price)`
  - Unique Zipcodes: `r n_distinct(NYC_Rental_Price$zipcode)`

- `NYC_Zipcode` Dataset
  - Dimension: `r nrow(NYC_Zipcode)` x `r ncol(NYC_Zipcode)`
  - Unique Zipcodes: `r n_distinct(NYC_Zipcode$zipcode)`

- `US_Housing_Price` Dataset
  - Dimension `r nrow(US_Housing_Price)` x `r ncol(US_Housing_Price)`

- The merged `Merged` Dataset
  - Dimension `r nrow(Merged)` x `r ncol(Merged)`
  - Unique Zipcodes: `r n_distinct(Merged$zipcode)`
  - Unique Neighborhoods: `r n_distinct(Merged$neighborhood)`

The distinction between datasets `NYC_Rental_Price` and `NYC_Zipcode` in terms of the number of Zipcodes is probably due to    
1. There are non-unique zipcode present in the `NYC_Zipcode` dataset    
1. Many records of the `NYC_Rental_Price` were removed due to incomplete data   
1. The `NYC_Zipcode` dataset is not covering every corner within NYC so it is not as thorough as the zipcodes in the `NYC_Zipcode` dataset    

After identifying non-unique zipcodes, rectify the cleaned datasets to ensure `zipcode` is unique-identifer among all datasets.
```{r Ensure zipcode as unique-identifier}
  
  NYC_Rental_Price %>% group_by(zipcode) %>% filter(n() > 1) %>% distinct(zipcode)  # zipcodes all unique
  NYC_Zipcode %>% group_by(zipcode) %>% filter(n() > 1) %>% distinct(zipcode)       # zipcode 10463 and 11201
  US_Housing_Price %>% group_by(zipcode) %>% filter(n() > 1) %>% distinct(zipcode)  # zipcodes all unique
  Merged %>% group_by(zipcode) %>% filter(n() > 1) %>% distinct(zipcode)            # zipcode 11201

  NYC_Zipcode %>% filter(zipcode==11201 | zipcode==10463) %>% select(zipcode, county, borough) # correct data should be: 11201--Kings--Brooklyn & 10463--Bronx--Bronx
  Merged %>% filter(zipcode==11201) %>% select(zipcode, county, borough)                       # correct data should be: 11201--Kings--Brooklyn
  
  NYC_Zipcode = NYC_Zipcode %>% filter(!(zipcode==11201 & county=="New York")) %>% filter(!(zipcode==10463 & county=="New York"))
  Merged = Merged %>% filter(!(zipcode==11201 & county=="New York"))
  
```


##### [Code Chunk,Table1] showing average rental price in each borough and year    

```{r pivot-longer table for datasets}

  Merged_Longer = pivot_longer(Merged, cols=c(5:ncol(Merged)), names_to="date", values_to="rental_price") %>% separate(col=date, into=c("year", "month", "day"), sep="/")
  NYC_Rental_Price_Longer = pivot_longer(NYC_Rental_Price, cols=c(3:ncol(NYC_Rental_Price)), names_to="date", values_to="rental_price") %>% separate(col=date, into=c("year", "month", "day"), sep="/")
  US_Housing_Price_Longer = pivot_longer(US_Housing_Price, cols=c(3:ncol(US_Housing_Price)), names_to="date", values_to="housing_price") %>% separate(col=date, into=c("year", "month", "day"), sep="/")
  
```

```{r Table showing average rental price in each borough and year}

  temp = Merged_Longer %>% group_by(borough, year) %>% summarise(mean_rental_price=mean(rental_price, rm.na=TRUE))
  Table1 = temp |> pivot_wider(names_from=borough, values_from=mean_rental_price) |> janitor::clean_names()
  
```

##### [Discussion]    
- The rental prices of all boroughs in New York City reveal a clear INCREASING TREND over years 2015 to 2024.   
- Manhattan has MOST EXPENSIVE mean rental prices each year from 2015 to 2024 among all five boroughs of New York City.
- Rental prices for Statan Island and Bronx are UNAVAILABLE.

##### [Code Chunk,Table2] comparing rental prices from 2020 to 2021
```{r Table comparing rental price from 2020 to 2021}
    
  temp = Merged_Longer %>% filter( (year=="2020"|year=="2021") & month==c("01") ) %>% group_by(year) %>% select(zipcode, year, rental_price) %>% rename("January_of_Year"="year")

  Table2 = temp |> pivot_wider(names_from=January_of_Year, values_from=rental_price) |> rename("Jan.2020"="2020", "Jan.2021"="2021") |> arrange(zipcode) |>
           mutate(absolute_change = Jan.2021-Jan.2020, relative_percentage_change = (Jan.2021 - Jan.2020) / Jan.2020 * 100) %>% mutate(across(where(is.numeric), \(x) round(x, 2)))
```

##### [Discussion]    
- Rent prices experience decrease for the majority of the zipcodes from 2020 to 2021.   
- Rental prices have fallen in 2021 relative to 2020 with an average drop ranging between 5% and 20%

##### [Code Chunk,Table3] showing largest drop from 2020 to 2021; include the neighborhood with the largest drop.
```{r Table showing largest drop from 2020 to 2021}
 
  temp = Merged_Longer %>% filter(year=="2020"|year=="2021") %>% select(borough, neighborhood, year, month, day, rental_price) %>% group_by(year, borough, neighborhood) %>%
         summarise(mean_rental_price=mean(rental_price)) %>% mutate(mean_rental_price = round(mean_rental_price, digits=2))
  Table3 = pivot_wider(temp, names_from=year, values_from=mean_rental_price) %>% mutate(absolute_change=`2021`-`2020`) %>% arrange(borough, absolute_change) %>% group_by(borough) %>%
           slice(1) %>% ungroup()

```

##### [Discussion]    
- Manhattan's East Harlem shows a significant reduction in rental prices compared to other areas, suggesting that it was particularly affected in period from 2020 to 2021.
- Brooklyn’s East New York and New Lots also saw a considerable decrease.
- Queens borough experienced relatively smaller changes.


##### [Code Chunk,Plot1] showing NYC Rental Prices within Zipcodes for all available years.


##### [Code Chunk,Plot2] comparison of distribution of house prices across states.
```{r Table4 & Plot2, fig.width=10, fig.height=10}

  temp = US_Housing_Price_Longer %>% group_by(zipcode, month, state) %>% summarise(mean_housing_price=mean(housing_price), .groups="keep") %>% 
         mutate(month=as.numeric(month), mean_housing_price=round(mean_housing_price,digits=0))

  Table4 = temp %>% pivot_wider(names_from = month, values_from = mean_housing_price)
           colnames(Table4)[3:ncol(Table4)] <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

  Plot2= ggplot(temp, aes(x=state, y=mean_housing_price, fill=state)) + geom_boxplot() + 
         labs(title = "Distribution of Zipcode-level House Prices Across States", x = "State", y = "House Price") +
         theme_light() + theme(plot.title = element_text(hjust = 0.5)) + theme(legend.position = "bottom") + guides(fill = guide_legend(ncol = 20)) + 
         theme(legend.key.size = unit(0.5, "cm")) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
  Plot2
         
```

##### [Discussion]    
- There is significant variation of house prices across states in 2023, as indicated by the wide range in the box heights and the presence of outliers.   
  For instance, states like California (CA) and New York (NY) show much higher house prices compared to states such as Kansas (KS) and Mississippi (MS).
- Many states have several outliers, particularly California (CA), New York (NY), and Florida (FL). These outliers represent exceptionally high house prices in certain zip 
  codes, possibly in premium or luxury neighborhoods.   
  Conversely, in some states (e.g., Wyoming (WY), North Dakota (ND)), the house prices are more clustered, with fewer extreme outliers.
- States like California (CA), New York (NY), and Hawaii (HI) generally have higher house prices, with their median values above $2 million in many cases.    
  On the other end of the spectrum, states like Arkansas (AR) and Mississippi (MS) have much lower house prices, with most prices well under $500,000.

##### [Code Chunk,Plot3] zipcode-specific housing prices against zipcode specific rental prices.
```{r Plot3, fig.width=10, fig.height=10}
  
  temp = NYC_Rental_Price_Longer %>% filter(year=="2023") %>% select(zipcode, month, rental_price) %>% group_by(zipcode) %>% summarise(annual_mean_rental=mean(rental_price))
  temp1= US_Housing_Price_Longer %>% filter(year=="2023") %>% select(zipcode, month, housing_price) %>% group_by(zipcode) %>% summarise(annual_mean_housing=mean(housing_price))
  temp2= left_join(x=temp, y=temp1, by="zipcode") %>% drop_na()

  Plot3_1 = ggplot(temp2, aes(x=as.factor(zipcode), y=annual_mean_rental, group=1)) + geom_line(color="darkgreen") +
            labs(title = "Annual Mean Rental Prices by Zipcode in 2023",x = "Zipcode", y = "Annual Mean Rental Price") + theme_light() +
            theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title=element_text(hjust=0.5))
  Plot3_2 = ggplot(temp2, aes(x=as.factor(zipcode), y=annual_mean_housing, group=1)) + geom_line(color="darkblue") +
            labs(title = "Annual Mean Housing Prices by Zipcode in 2023", x = "Zipcode", y = "Annual Mean Housing Price") + theme_light() +
            theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title=element_text(hjust=0.5))

  Plot3_1 / Plot3_2

```

##### [Discussion]    
- We can see that the Housing Prices are overall significantly higher than the Rental Prices.   
- The rental prices generally seem to cluster around the $3,000 to $4,500 range.    
  On the other hand, the disparity between the most expensive and least expensive areas in terms of housing prices is substantial.    
- Some zip codes, such as 10012 and 10013, stand out with higher rental prices that exceed $5,000.    
  Certain zip codes like 10012 and 10013 have significantly higher housing prices, exceeding $3,000,000.    
- The variation in housing prices is much larger compared to rental prices.     













